---
title: "Example Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Analysis}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
library(gganimate)
library(tidyverse)
library(lubridate)
```


## Intruduction

This vignette demonstrates the functions of the gganimate package using a real-world dataset. The dataset consists of the daily stock prices and volume of 14 different tech companies, including Apple (AAPL), Amazon (AMZN), Alphabet (GOOGL), and Meta Platforms (META) and more! The data was obtained from [this link](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-02-07/readme.md).

The questions I aim to answer: How did the stock prices of five tech companies (Apple, Amazon, Alphabet, Meta, and Microsoft) change over time?

## Data source and dictionary

The original data come from [Yahoo Finance](https://stackoverflow.com/questions/56024713/how-to-use-gganimate-package-plot-a-dynamic-map-in-r), which provides historical stock data for various companies. The data was scraped and cleaned by [Ryo Nakagawara](https://anderfernandez.com/en/blog/how-to-create-animations-in-r-with-gganimate/) and posted on [Kaggle](https://www.kaggle.com/datasets/evangower/big-tech-stock-prices/).

The data dictionary for the dataset is as follows:

| Variable | Description |
| -------- | ----------- |
| stock_symbol | stock_symbol |
| date | date |
| open | The price at market open. |
| high | The highest price for that day. |
| low | The lowest price for that day. |
| close | The price at market close, adjusted for splits. |
| adj_close |	double |	The closing price after adjustments for all applicable splits and dividend distributions. Data is adjusted using appropriate split and dividend multipliers, adhering to Center for Research in Security Prices (CRSP) standards. |
| volume | The number of shares traded on that day. |

```{r}
tuesdata <- tidytuesdayR::tt_load('2023-02-07')
tuesdata <- tidytuesdayR::tt_load(2023, week = 6)

big_tech_stock_prices <- tuesdata$big_tech_stock_prices
big_tech_companies <- tuesdata$big_tech_companies
```
## Data analysis

### Filter the stocks data to keep only the five companies of interest

```{r}
big_tech_stock_prices <- big_tech_stock_prices %>%
  filter(stock_symbol %in% c("AAPL", "AMZN", "GOOGL", "MSFT", "TSLA"))
```

```{r}
big_tech_stock_prices <- big_tech_stock_prices %>%
  mutate(date = as.Date(date))
```

### Compute the percentage change in stock price from the previous day

```{r}
big_tech_stock_prices <- big_tech_stock_prices %>%
  group_by(stock_symbol) %>%
  mutate(pct_change = (close - lag(close)) / lag(close) * 100) %>%
  ungroup()
```

### Plot the stock price over time for each company

```{r}
p1 <- ggplot(big_tech_stock_prices, aes(x = date, y = close, color = stock_symbol)) +
  geom_line() +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Stock price over time",
       subtitle = "Data source: Yahoo Finance via Kaggle",
       x = "Date",
       y = "Stock price (USD)",
       color = "Company")
print(p1)
```

### Plot the percentage change in stock price over time for each company

```{r}
p2 <- ggplot(big_tech_stock_prices, aes(x = date, y = pct_change, color = stock_symbol)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Percentage change in stock price over time",
       subtitle = "Data source: Yahoo Finance via Kaggle",
       x = "Date",
       y = "Percentage change",
       color = "Company")
print(p2)
```

### Plot the stock price distribution for each company by year

```{r}
p3 <- ggplot(big_tech_stock_prices, aes(x = close, fill = stock_symbol)) +
  geom_density(alpha = 0.7) +
  scale_x_continuous(labels = scales::dollar_format()) +
  facet_wrap(~year(date)) +
  labs(title = "Stock price distribution by year",
       subtitle = "Data source: Yahoo Finance via Kaggle",
       x = "Stock price (USD)",
       y = "Density",
       fill = "Company")
print(p3)
```

```{r}
# Create a point plot of closing price vs volume for each company, faceted by sector
big_tech_stock_prices <- big_tech_stock_prices %>%
  mutate(sector = case_when(
    stock_symbol %in% c("AAPL", "MSFT") ~ "Technology",
    stock_symbol %in% c("AMZN", "GOOGL") ~ "Communication Services",
    stock_symbol == "TSLA" ~ "Consumer Discretionary"
  ))
p4 <- ggplot(big_tech_stock_prices, aes(x = volume, y = close, color = stock_symbol)) +
  geom_point() +
  scale_x_log10(labels = scales::comma_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(title = "Closing price vs volume of big tech stocks as of 2023-02-07",
       subtitle = "Data from Kaggle",
       x = "Volume (log scale)",
       y = "Closing price (USD)",
       color = "Company") +
  facet_wrap(~ sector)

print(p4)
```

### Apply gganimate functions
```{r}
p1_animated <- p1 +
  transition_reveal(date) +
  enter_fade() +
  exit_fade() +
  ease_aes("linear")

p2_animated <- p2 +
  transition_reveal(date) +
  enter_fade() +
  exit_fade() +
  ease_aes("linear")
```

### Summarize

The plots show that the stock prices of the five big tech companies increased over time, with some fluctuations and drops. After 2020, the stock prices of these companies have increased significantly.

### List the functions used

dplyr: filter, mutate, group_by, ungroup, lag
lubridate: as.Date, year
ggplot, aes, geom_line, scale_y_continuous, scale_x_continuous, labs, geom_density, facet_wrap
gganimate: transition_reveal, enter_fade, exit_fade, ease_aes
